---
title: "Data Analysis"
date: "August 2023"
author: "Catherine Cornella"
output: 
  html_document:
    code_folding: show
---

# Setup libraries

The packages required are as follows: 

* tidyverse (for dplyr, readr, tidyr)
* stringr (for stringr)
* knitr (for knitr)
* lubridate (for lubridate)
* ggplot2 (for ggplot2)

```{r Setup:Libraries, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)      #General Data wrangling
library(ggplot2)    #Plotting and analysis
library(readr)      #File reading and writing
library(stringr)    #String reading
library(knitr)      #Markdown file exporting
library(tidyr)      #Data formatting
library(lubridate)  #Date-Time Wrangling
```


```{r Setup:Variables, message=FALSE, warning=FALSE}

#Establish General Import path: 
#If the data location is ever changed, it can be altered in this one location and hold true for every file. 
importPath="/data/airflow/analysis/csvs/"

#Start date: 
startDate <- "2022-08-18"
endDate <- "2023-04-16"
```

All files are expected to be under the import path given above, under `csvs`. Used repeatedly throughout the analysis is also the window of time that all data is narrowed to. We specify that here so as to change more easily. 

# Import data

### Meteorological data

```{r Import:Meteor, message=FALSE, warning=FALSE}
path=paste(importPath, "mkws_data.csv", sep="")
MeteorDataDT <- read_csv(path, skip=2)
```

### Sensor Data

The sensor files names will have to be changed when their locations changed. This is something of a hardcoded bug, but it is on purpose. It could be made to simply select all sensor data, but that carries risks of mixing the data from sensors now in different locations. At present, simply changing the names of the csvs below offers the most control. 

```{r Import:Sensor, message=FALSE, warning=FALSE}
path=paste(importPath, "cyclone-hx9-telescope_spider_south.csv", sep="")
cyc_df <- read_csv(path, skip=5)
path=paste(importPath, "darkthunder-dome_slit_bottom_horizontal.csv", sep="")
dark_df <- read_csv(path, skip=5)
path=paste(importPath, "freflow-mez_obsroom_horizontal.csv", sep="")
fre_df <- read_csv(path, skip=5)
path=paste(importPath, "picasso-mez_obsroom_vertical.csv", sep="")
pic_df <- read_csv(path, skip=5)
path=paste(importPath, "turbopanda-dome_slit_top_vertical.csv", sep="")
turbo_df <- read_csv(path, skip=5)
```

### Engineering Data

The engineering data is in one file, engineeringMaster.csv, expected output of `csv_to_master.R`. 

```{r Import:Engineering, message=FALSE, warning=FALSE}
path=paste(importPath, "engineeringMaster.csv", sep="")
engineeringMaster <- read_csv(path, skip=3)
```

# Tidy/Re-format Data

To this point, the data has been brought over with as minimal manipulation or exclusion of data as possible. However, it is easier to run analysis on data with standardized dates and times, and not necessary to keep weather data beyond our window of time. 

The weather data comes as just all the data from a particular year, so we narrow it down to our specific time window. 

```{r Tidy:Meteor, message=FALSE, warning=FALSE}
#Meteorlogical data: 
#Narrow the dates down to our timeframe. 
weather_d <- MeteorDataDT %>% filter(MeteorDataDT$DateTime >= as.Date(startDate) & MeteorDataDT$DateTime <= as.Date(endDate) )
```

The sensor data is also combined into a single master dataframe. 

```{r Tidy:Sensor, message=FALSE, warning=FALSE}
#Add all the Sensors into one dataframe.
sensorMaster <- rbind(cyc_df, dark_df, fre_df, pic_df, turbo_df)
```

# Analysis

Exploratory graph

```{r message=FALSE, warning=FALSE}
#Cn2 taken as log for readability. 
#Simple Cn2 values by sensor, big picture, binwidth approach
sensorMaster %>%  
ggplot( aes(x=(log(Cn2, base=10)), color=sensor)) + 
  geom_histogram(binwidth=0.25)+ 
  labs(title="Turbulence Value Distribution", x= "log 10 of Cn2 values", y= "count of values")
 
#Simple Cn2 values by sensor, big picture, bins approach
sensorMaster %>%  
ggplot( aes(x=(log(Cn2, base =10)), color=sensor)) + 
  geom_histogram(bins=100)+ 
  labs(title="Turbulence Value Distribution (Percentile)", x= "log 10 of Cn2 values", y= "count of values")

#Simple Cn2 values by sensor, facetwrapped, 5xbins
sensorMaster %>% 
ggplot( aes(x=(log(Cn2, base=10)), color=sensor)) + 
  geom_histogram(bins=500) +
  facet_wrap(~sensor, scales="fixed")+ 
  labs(title="Turbulence Value Distribution by sensor", x= "log 10 of Cn2 values", y= "count of values")
```
```{r}
#Simple Cn2 values by sensor, facet~wrapped by dome open or closed

#Slit legend: 
# 0: dome midway
# 1: dome open
# -1: dome closed
# -2: moving

# try to mutate in the dome slit naively and see if it works (spoilers, seems to)
sensorSlit <- engineeringMaster %>% select(dateTime, slit) %>%
                                    mutate(slit=round(slit)) %>%
                                    merge(sensorMaster, by="dateTime", all=TRUE) %>% 
                                    drop_na(sensor)


# Show by sensor when dome is Open
sensorSlit %>% filter(slit=="0") %>% 
ggplot( aes(x=(log(Cn2, base=10)), color=sensor)) +
  geom_histogram(bins=500) +
  facet_wrap(~sensor, scales="fixed") + 
  labs(title="Dome open by sensor", x= "log 10 of Cn2 values", y= "count of values")

# Show by sensor when dome is closed
sensorSlit %>% filter(slit=="-1") %>% 
ggplot( aes(x=(log(Cn2, base = 10)), color=sensor)) +
  geom_histogram(bins=500) +
  facet_wrap(~sensor, scales="fixed") + 
  labs(title="Dome closed by sensor", x= "log 10 of Cn2 values", y= "count of values")

```


(1) What is the average daytime and nighttime (or maybe nighttime+dome open) Cn2 values per sensor?
What are the corresponding standard deviations?  

```{r message=FALSE, warning=FALSE}
#make the dataframe: average Cn2 values per sensor, split between day and night, by hour. 

# Okay. The point here is to make the average values of the Cn2 values every hour. 
# So we first group by the sensor. this means any like sum thing we do won't be all of them, but will isolate them based on sensor 
# Then we create a new column with just the hour because we don't really care about the minutes of the night, and it was proving too difficult to do so. 
# Then we once again group by sensor, but also by hourTime, so that
# we can then take the average Cn2 value, and it will give us the average cn2 value at that time of night from that sensor. 

day_avg_byHour <- sensorMaster %>% group_by(sensor) %>% 
                            mutate(hourTime= hour(dateTime)) %>%
                            filter(hourTime >= 5 & hourTime <= 18) %>% 
                            group_by(sensor, hourTime) %>% 
                            mutate(std_dev =sd(Cn2)) %>% 
                            group_by(sensor, hourTime, std_dev) %>% 
                            summarize(hourAvg= mean(Cn2))

#plot
day_avg_byHour_plot <- day_avg_byHour %>% ggplot(aes(x=hourTime, color=sensor)) +
                            geom_line(aes(y=hourAvg )) + 
                            geom_line(aes(y=std_dev), linetype="dashed") + 
                            facet_wrap(~sensor, scales="fixed")+ 
                            labs(title="Average Cn2 values (solid) by sensor with standard deviation (dashed)", 
                                 x= "Hour of Day", y= "Cn2 Values (fixed)")
#display
day_avg_byHour_plot

day_avg_byHour_plot <- day_avg_byHour %>% ggplot(aes(x=hourTime, color=sensor)) +
                            geom_line(aes(y=hourAvg )) + 
                            geom_line(aes(y=std_dev), linetype="dashed") + 
                            facet_wrap(~sensor, scales="free")+ 
                            labs(title="Average Cn2 values (solid) by sensor with standard deviation (dashed)", 
                                 x= "Hour of Day", y= "Cn2 Values (free)")
#display
day_avg_byHour_plot

```

I note that everything so far has been seemingly ignoring the time discrepancies between all the sensors. 

Further Questions to explore: 

Do these values change if you cut your sample size in half (first half and second half or odd samples v. even samples)?  
For nighttime dome open measurements (and maybe separately for closed dome measurements)  

(1) do the sensor values show any dependence on difference in temperature between the 
       a) mirror and the tube, 
or the b) mirror and the outside air,
or the c) tube and the outside air? 

(2) versus time of day, and 
(3) versus the direction of the dome with respect to the wind direction reported by the CFHT/Gemini weather tower?
